---
output:
  html_document: default
  pdf_document: default
---

# Data Mining Lab 3 - Classification

---
Title: Data Mining Lab 3
Authors: Max Link, Logan Lu, Jadon Klipsch
Date: "2025-04-14"
Description: In this project, we will focus on classification 
---

```{r setup, include=FALSE}
# Disable RStudio's package update check for this session
options("rstudio.package_manager.check_updates" = FALSE)

# Set the CRAN repository
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install packages quietly
if (!requireNamespace("xfun", quietly = TRUE)) {
  install.packages("xfun", quietly = TRUE)
}
if (!requireNamespace("ggrepel", quietly = TRUE)) {
  install.packages("ggrepel", quietly = TRUE)
}

# Set knitr options
knitr::opts_chunk$set(echo = TRUE)

# Load required libraries and suppress startup messages
suppressPackageStartupMessages({
  library(dplyr)      # For data manipulation
  library(ggplot2)    # For visualizations
  library(tidyr)      # For cleaning data
  library(ggrepel)    # For repelling labels in plots
})
```

```{r, results='hide', message=FALSE, warning=FALSE}
# add packages to install here
pkgs <- c("tidyverse", "factoextra", "cluster", "patchwork", "tibble", "ggrepel",
          "mclust", "mcclust", "fpc", "seriation", "apcluster", "dbscan", "entropy", "maps", "kernlab", "skimr")
pkgs_install <- pkgs[!(pkgs %in% installed.packages()[,"Package"])]
if(length(pkgs_install)) install.packages(pkgs_install)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(mclust))
suppressPackageStartupMessages(library(mcclust))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(fpc))
suppressPackageStartupMessages(library(seriation))
suppressPackageStartupMessages(library(apcluster))
suppressPackageStartupMessages(library(dbscan))
suppressPackageStartupMessages(library(entropy))
suppressPackageStartupMessages(library(maps))
suppressPackageStartupMessages(library(skimr))
``` 

```{r, results='hide', message=FALSE, warning=FALSE}


# Load libraries
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(tibble)
library(mclust)
library(mcclust)
library(ggplot2)
library(ggrepel)
library(cluster)
library(patchwork)
library(factoextra)
library(fpc)
library(seriation)
library(apcluster)
library(dbscan)
library(entropy)
library(maps)
library(skimr)


```

```{r}

dataset <- read.csv("tx_data_with_gr.csv")

# print(dataset)
```

### Section 1 -- Data Prep

First, we need to define our classes.

We can pick out two classes: 

1) High vs Low Infection Rate 



2) High vs Low Death Rate 



Now we can explore statistics to get proper thresholds

```{r}
# 1) Explore distribution of infection_rate and death_rate
# Summary statistics
cat("Summary of Infection Rate:\n")
summary(dataset$infection_rate)
cat("\nSummary of Death Rate:\n")
summary(dataset$death_rate)

```
The median is appropriate for both of these classification tasks because choosing the median balances our classes. Dealing with balanced classes means that we have a close 50-50 county split into high and low categories. This even split enables logistic regression, decision trees, SVMs. 

The infection rate threshold should be 0.07212. 

The death rate threshold should be 0.022603. 

```{r}
# Load required library (optional, for table creation)
library(dplyr)

# Read the dataset (update path as needed)
dataset <- read.csv("tx_data_with_gr.csv", stringsAsFactors = FALSE)

# Total number of counties
total_counties <- nrow(dataset)

# 1) High vs. Low Infection Rate (threshold: 0.1)
# Count counties with infection_rate > 0.1 (High)
high_infection <- sum(dataset$infection_rate > 0.07212, na.rm = TRUE)
# Count counties with infection_rate <= 0.1 (Low)
low_infection <- sum(dataset$infection_rate <= 0.07212, na.rm = TRUE)
# Calculate percentages
high_infection_pct <- (high_infection / total_counties) * 100
low_infection_pct <- (low_infection / total_counties) * 100

# 2) High vs. Low Death Rate (threshold: 0.023, updated from 0.0023)
# Count counties with death_rate > 0.023 (High)
high_death <- sum(dataset$death_rate > 0.022603, na.rm = TRUE)
# Count counties with death_rate <= 0.023 (Low)
low_death <- sum(dataset$death_rate <= 0.022603, na.rm = TRUE)
# Calculate percentages
high_death_pct <- (high_death / total_counties) * 100
low_death_pct <- (low_death / total_counties) * 100

# Print results
cat("1) Infection Rate Classification (Threshold: 0.07212, i.e., 1,000 cases per 10,000 population)\n")
cat("High Infection Rate (> 0.07212):", high_infection, "counties (", round(high_infection_pct, 2), "%)\n")
cat("Low Infection Rate (<= 0.07212):", low_infection, "counties (", round(low_infection_pct, 2), "%)\n\n")

cat("2) Death Rate Classification (Threshold: 0.022603, i.e., 230 deaths per 10,000 population)\n")
cat("High Death Rate (> 0.022603):", high_death, "counties (", round(high_death_pct, 2), "%)\n")
cat("Low Death Rate (<= 0.022603):", low_death, "counties (", round(low_death_pct, 2), "%)\n\n")

```

Now we can make our table with our class attributes for infection rate classification and death rate classification. 

```{r}
# Create class attributes for Infection Rate and Death Rate
dataset <- dataset %>%
  mutate(Infection_Rate_Class = ifelse(infection_rate > 0.07212, "High", "Low"),
         Death_Rate_Class = ifelse(death_rate > 0.022603, "High", "Low"))

# Select features for model learning, including both class attributes
table_for_model <- dataset %>%
  select(county_name, median_age, median_income, total_pop,
         rate_of_growth_first_30_days, rate_of_growth_30_60_days,
         rate_of_growth_60_90_days, rate_of_growth_90_120_days,
         rate_of_growth_120_150_days, rate_of_growth_150_180_days,
         Infection_Rate_Class, Death_Rate_Class)

# Display first 10 rows
cat("\nSingle Table with Infection_Rate_Class and Death_Rate_Class (First 10 Rows):\n")
print(head(table_for_model, 10))

# Verify class distributions
# Infection Rate
infection_dist <- table(dataset$Infection_Rate_Class)
infection_prop <- prop.table(infection_dist)
cat("\nInfection Rate Class Distribution:\n")
print(infection_dist)
cat("\nInfection Rate Class Proportions:\n")
print(infection_prop)

# Death Rate
death_dist <- table(dataset$Death_Rate_Class)
death_prop <- prop.table(death_dist)
cat("\nDeath Rate Class Distribution:\n")
print(death_dist)
cat("\nDeath Rate Class Proportions:\n")
print(death_prop)

# Save the table
write.csv(table_for_model, "tx_data_with_class.csv", row.names = FALSE)

```
